---
title: "The bodymapRat data user's guide"
author: 
- name: Stephanie Hicks
    affiliation: Johns Hopkins Bloomberg School of Public Health
    email: shicks19@jhu.edu
- name: Kwame Okrah
    affiliation: Genentech
    email: kwame.okrah@gmail.com
date: Feburary 05, 2019
package: bodymapRat
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

<!--
    %% \VignetteEngine{knitr::knitr}
-->

## Introduction

The `bodymapRat` package contains gene expression 
data on 652 RNA-Seq samples from a comprehensive rat 
transcriptomic BodyMap study. These samples include 
the sequence identifier information provided in the 
header of the FASTQ files which can be used as a 
surrogate for batch. These samples have not been
normalized or pre-processed.

The data are provided in a `SummarizedExperiment`. The 
phenotypic information can be extracted using the 
`colData()` function and a description of the phenotypic
data is listed in the table below:

| Title | Description |
| :--- | :--- |
| sraExperiment | SRA Experiment ID | 
| title | Title of sample provided by the authors | 
| geoAccession | GEO Accession ID | 
| BioSample | BioSample ID | 
| avgLength | Average read length | 
| instrument | Machine identifier (from FASTQ header) | 
| runID | Run ID (from FASTQ header) | 
| fcID | Flow cell ID (from FASTQ header) | 
| fcLane | Flow cell lane (from FASTQ header) | 
| tile | Tile (from FASTQ header) | 
| xtile | xtile (from FASTQ header) | 
| ytile | ytile (from FASTQ header) | 
| organ | Body organ | 
| sex | Gender | 
| stage | Stage | 
| techRep | Technical replicate number | 
| colOrgan | Column of colors to help with plotting | 
| rnaRIN | RIN number | 
| barcode | barcode number | 


The data can be accessed as follows:

```{r loadlibs, message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(bodymapRat)
```

```{r loaddata}
data(bodymapRat)

# Get the expression data
counts = assay(bodymapRat)
dim(counts)
counts[1:5, 1:5]

# Get the pheno data
head(colData(bodymapRat))
```

The data in this package are used as an example
data set in the `qsmooth` package.

# References

1. Yu et al. (2013). A rat RNA-Seq transcriptomic
BodyMap across 11 organs and 4 developmental stages.
_Nature Communications_ **5**:3230. PMID: 24510058.
PMCID: PMC3926002.


# SessionInfo

```{r sessionInfo,results ='markup'}
sessionInfo()
```


