---
title: "The bodymapRat data user's guide"
author:
- name: Stephanie C. Hicks
  affiliation: Johns Hopkins Bloomberg School of Public Health
- name: Kwame Okrah
  affiliation: Genentech
output:
  BiocStyle::html_document:
    toc_float: true
package: bodymapRat
abstract: |
 This package contains a SummarizedExperiment
   from the Yu et al. (2013) paper that performed the 
   rat BodyMap across 11 organs and 4 developmental stages. 
   Raw FASTQ files were downloaded and mapped using STAR. 
   Data is available on ExperimentHub as a data package.
vignette: |
  %\VignetteIndexEntry{The bodymapRat data user's guide}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, echo=FALSE, results="hide", message=FALSE}
require(knitr)
opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
```

```{r style, echo=FALSE, results='asis'}
BiocStyle::markdown()
```

## Introduction

The `bodymapRat` package contains gene expression 
data on 652 RNA-Seq samples from a comprehensive rat 
transcriptomic BodyMap study. These samples include 
the sequence identifier information provided in the 
header of the FASTQ files which can be used as a 
surrogate for batch. These samples have not been
normalized or pre-processed.

The data are provided in a `SummarizedExperiment`. The 
phenotypic information can be extracted using the 
`colData()` function and a description of the phenotypic
data is listed in the table below:

| Title | Description |
| :--- | :--- |
| sraExperiment | SRA Experiment ID | 
| title | Title of sample provided by the authors | 
| geoAccession | GEO Accession ID | 
| BioSample | BioSample ID | 
| avgLength | Average read length | 
| instrument | Machine identifier (from FASTQ header) | 
| runID | Run ID (from FASTQ header) | 
| fcID | Flow cell ID (from FASTQ header) | 
| fcLane | Flow cell lane (from FASTQ header) | 
| tile | Tile (from FASTQ header) | 
| xtile | xtile (from FASTQ header) | 
| ytile | ytile (from FASTQ header) | 
| organ | Body organ | 
| sex | Gender | 
| stage | Stage | 
| techRep | Technical replicate number | 
| colOrgan | Column of colors to help with plotting | 
| rnaRIN | RIN number | 
| barcode | barcode number | 


The data can be accessed as follows:

```{r loadlibs, message=FALSE, warning=FALSE}
library(SummarizedExperiment) 
library(bodymapRat)
```

We use the `bodymapRat()` function to download the 
relevant files from _Bioconductor_'s 
[ExperimentHub](http://bioconductor.org/packages/ExperimentHub) web 
resource. Running this function will download a 
`SummarizedExperiment` object, which contains read counts, as 
well as the metadata on the rows (genes) and columns (cells). 

```{r loaddata}
bm_rat <- bodymapRat()

# Get the expression data
counts = assay(bm_rat)
dim(counts)
counts[1:5, 1:5]

# Get the meta data along columns
head(colData(bm_rat))
```

The data in this package are used as an example
data set in the 
[`qsmooth` Bioconductor package](http://bioconductor.org/packages/qsmooth).

# References

1. Yu et al. (2013). A rat RNA-Seq transcriptomic
BodyMap across 11 organs and 4 developmental stages.
_Nature Communications_ **5**:3230. PMID: 24510058.
PMCID: PMC3926002.


# SessionInfo

```{r sessionInfo}
sessionInfo()
```


